### 2021-07-29 doc/thesis/chapters/deployment-on-hardware.tex
fix all the 'Gedankenstriche' in sentences

The CPU is the main actor that determines any actions on the system.
First, it reads a file from the external SD card and stores this data in a dedicated area in the memory.
Next, the DMA is configured and its data transfer mode is initiated, so that the DMA starts to transfer data from the mentioned memory area towards the FPGA.
The DMA is configured in a mode that creates a continuous transfer, by looping through the data - once the end is reached, the transfer starts at the beginning of the memory area.
Any data connection in the following processing chain is implemented as an AXI stream bus.
The first block after the DMA is an AXI Stream Switch.
It can switch its single input to be forwarded to either of the outputs.
The switch is configured by the CPU, to enable one output at a time.
Consequently, one of the FM radio receiver IPs is provided with data.
The IP processes this data to produce an audio output, which is delivered to another AXI Stream Switch.
It fulfills the function to either select the output of above or below IP, in coordination with the first switch, obviously.
The next block is an AXI Stream Broadcaster, which takes a single stream input and simply duplicates it at its output, so that the outputs are an exact copy of the input stream.
One of these duplicates is sourced to an AXI-Stream-To-I2S converter, which generates the I2S signals for the external audio codec chip.
The second duplicate is connected to a FIFO, which simply stores the generated output data.
Once the FIFO is almost full, it asserts an interrupt to the CPU, which triggers it to read out the entire FIFO buffer.
The CPU then stores this data on the external SD card again.
In that way, the loop of data is closed, which enables the verification of the respective IPs' correct functionality.
Therefore, this system design can be described as a hardware-in-the-loop system, in the point of view of the FPGA IPs.
This ability refers to the knowledge gained with the SIMILAR concept, that is introduced in Section \ref{sec:SimilarProcess}.
Specifically, this implemented system design enables the application of the steps 'Assess Performance' and 'Re-evaluate', since the IPs' performance can be measured and their design can be re-evaluated.\\

>>>>

The CPU is the main actor that determines any actions on the system.
First, it reads a file from the external SD card and stores this data in a dedicated area in the memory.
Next, the DMA is configured and its data transfer mode is initiated, so that the DMA starts to transfer data from the mentioned memory area towards the FPGA.
The DMA is configured in a mode that creates a continuous transfer, by looping through the data -- once the end is reached, the transfer starts at the beginning of the memory area.
Any data connection in the following processing chain is implemented as an AXI stream bus.
The first block after the DMA is an AXI Stream Switch.
It can switch its single input to be forwarded to either of the outputs.
The switch is configured by the CPU, to enable one output at a time.
Consequently, one of the FM radio receiver IPs is provided with data.
The IP processes this data to produce an audio output, which is delivered to another AXI Stream Switch.
It fulfills the function to either select the output of above or below IP, in coordination with the first switch, obviously.
The next block is an AXI Stream Broadcaster, which takes a single stream input and simply duplicates it at its output, so that the outputs are an exact copy of the input stream.
One of these duplicates is sourced to an AXI-Stream-To-I2S converter, which generates the I2S signals for the external audio codec chip.
The second duplicate is connected to a FIFO, which simply stores the generated output data.
Once the FIFO is almost full, it asserts an interrupt to the CPU, which triggers it to read out the entire FIFO buffer.
The CPU then stores this data on the external SD card again.
In that way, the loop of data is closed, which enables the verification of the respective IPs' correct functionality.
Therefore, this system design can be described as a hardware-in-the-loop system, in the point of view of the FPGA IPs.
This ability refers to the knowledge gained with the SIMILAR concept, that is introduced in Section \ref{sec:SimilarProcess}.
Specifically, this implemented system design enables the application of the steps 'Assess Performance' and 'Re-evaluate', since the IPs' performance can be measured and their design can be re-evaluated.\\

### 2021-07-29 doc/thesis/chapters/dsp.tex
fix all the 'Gedankenstriche' in sentences

Figure \ref{fig:fm_channel_baseband_freqs} shows the frequency spectrum of a single channel, located at its carrier- or center frequency.
For a better overview, only the positive spectum part is shown.
The negative part is an exact mirror, because the transmitted and received signals from the antenna are real signals - they do not have an imaginary part.
The spectrum consists of multiple parts.
Therefore, it is also referred to as `multiplex signal' in the literature.
Several parts are described in the following paragraphs \cite{EbuFmRegulators},\cite{EtsiFmSoundBroadcastingServiceHarmonizedStandard},\cite{ElectronicsNotesFmTutorial}.

@@

A frequency discriminator generates an output that is directly proportional to the frequency of the input FM signal.
In other words, it converts FM to AM.
The discriminator consists of two main parts - a differentiator and an envelope detector.
Different strategies can be implemented for both parts.
A block diagram for a frequency discriminator with its two main parts is shown in Figure \ref{fig:bd_freq_discriminator}.

@@

The demodulation of an FM channel to recover a stereo audio signal, requires a more sophisticated approach and multiple steps in signal processing.
Two parts need to be combined in order to achieve stereo.
The sum and the difference of the left and right audio channel signals - the mono and the stereo difference part, respectively.
The following equations need to be applied, to compute a left and a right channel.

>>>>

Figure \ref{fig:fm_channel_baseband_freqs} shows the frequency spectrum of a single channel, located at its carrier- or center frequency.
For a better overview, only the positive spectum part is shown.
The negative part is an exact mirror, because the transmitted and received signals from the antenna are real signals -- they do not have an imaginary part.
The spectrum consists of multiple parts.
Therefore, it is also referred to as `multiplex signal' in the literature.
Several parts are described in the following paragraphs \cite{EbuFmRegulators},\cite{EtsiFmSoundBroadcastingServiceHarmonizedStandard},\cite{ElectronicsNotesFmTutorial}.

@@

A frequency discriminator generates an output that is directly proportional to the frequency of the input FM signal.
In other words, it converts FM to AM.
The discriminator consists of two main parts -- a differentiator and an envelope detector.
Different strategies can be implemented for both parts.
A block diagram for a frequency discriminator with its two main parts is shown in Figure \ref{fig:bd_freq_discriminator}.

@@

The demodulation of an FM channel to recover a stereo audio signal, requires a more sophisticated approach and multiple steps in signal processing.
Two parts need to be combined in order to achieve stereo.
The sum and the difference of the left and right audio channel signals -- the mono and the stereo difference part, respectively.
The following equations need to be applied, to compute a left and a right channel.

### 2021-07-29 doc/thesis/chapters/hls.tex
check all \ref{} for a unified format
fix all the 'Gedankenstriche' in sentences

This chapter covers the High-Level Synthesis of \cplusplus\ in general, but also describes details that specifically apply to HLS tools that are provided by Xilinx.
An overview of the most important steps in the development workflow, as well as main features and possibilities of the code implementation and testbench are given.
The main sources of information for this chapter are found in \cite{VivadoUgHLSIntro}and \cite{VivadoUgHLS}.

@@

The usage of HLS tries to bring software- and hardware development closer together and tries to make these two worlds more similar.
In software development, there are different compilers for multiple processor architectures, which can be used to port a program to another target hardware.
The idea of HLS is to create the same level of abstraction - write code in a high-level language, which is agnostic of its target hardware architecture.\\

@@

The general significance of the simulation testbench for a HDL design is explained here, as well as the different levels of abstraction in which a HLS testbench can be executed.
These different levels of abstraction in the testbench execution are a major difference between a VHDL testbench and an HLS testbench.
The most important impact thereof is the execution speed - the time it takes to simulate a number of test cases on the DUT.\\

>>>>

This chapter covers the High-Level Synthesis of \cplusplus\ in general, but also describes details that specifically apply to HLS tools that are provided by Xilinx.
An overview of the most important steps in the development workflow, as well as main features and possibilities of the code implementation and testbench are given.
The main sources of information for this chapter are found in \cite{VivadoUgHLSIntro} and \cite{VivadoUgHLS}.

@@

The usage of HLS tries to bring software- and hardware development closer together and tries to make these two worlds more similar.
In software development, there are different compilers for multiple processor architectures, which can be used to port a program to another target hardware.
The idea of HLS is to create the same level of abstraction -- write code in a high-level language, which is agnostic of its target hardware architecture.\\

@@

The general significance of the simulation testbench for a HDL design is explained here, as well as the different levels of abstraction in which a HLS testbench can be executed.
These different levels of abstraction in the testbench execution are a major difference between a VHDL testbench and an HLS testbench.
The most important impact thereof is the execution speed -- the time it takes to simulate a number of test cases on the DUT.\\

### 2021-07-29 doc/thesis/chapters/implementation.tex
fix all the 'Gedankenstriche' in sentences

In the implemented version, an audio file is modulated into the FM channel.
The audio file contains data that represents speech with the words \textit{'Left channel... Right channel...'}.
In terms of audio channels, the words exactly match the respective channels' output - the first part is modulated on the left audio channel only, while the second part only targets the right audio channel.
This is particulary useful for verification, especially to verify the functionality of channel separation.
From the experience during development, the channel separation is a good indicator of whether the system is functioning correctly.
This left-right indication also simplifies the verification process, because a developer can recognize it with their visual and acoustic senses.
It is possible to audibly verify it by listening to the output.
It is also possible to just visually inspect the output data, where a clear separation is visible between the channels.\\

@@

In the final chain of DSP blocks, the sample rate is increased again, to be able to frequency-modulate the signal.
The resulting frequency modulated signal represents the signal that is received by an antenna - the RF signal of an FM channel.
This signal is then processed by a complex baseband mixer, which results in inphase-, and quadrature components in the baseband.
An LP filter band-limits the signal and removes artifacts of the mixer.
As a final step, the sample rate is decreased again, back to the chosen sampling rate of 960~kHz.
This signal can now be processed in a digital FM receiver.

@@

Again, an audio file is used as an input, which is modulated as an FM channel signal.
The entire multiplex signal is then modulated onto an actual carrier frequency of 99~MHz, that is within the FM radio broadcast band.
This entire DSP chain is being run in the GNU Radio software, on a host PC.
Only the final signal - FM modulated, and on a carrier of 99~MHz - is then sent to the USRP device in the form of digital signal samples.
The USRP hardware then transmits the signal via its antenna.\\

@@

The usage of a software like GNU Radio in combination with external hardware is an interesting tool for a use-case like this project.
An FM receiver is to be implemented in multiple different versions.
All of them need to be verified somehow - this is where the SDR transmitter created with GNU Radio and external hardware opens up an entire new possibility from the viewpoint of system level design integration.\\

@@

The testbench framework cocotb is developed as a free, open-source project.
It can be used to verify VHDL and SystemVerilog RTL code, in combination with an external simulator.
Various different simulators can be used with cocotb - amongst others, GHDL is supported.
The interface that is used to the simulator is the industry-standard VPI.\\

@@

The Python classes \textit{tb\_analyzer\_helper}, \textit{tb\_data\_handler} and \textit{fm\_receiver\_model} are implemented to create a better structure of the testbench and to provide code that is re-useable for the testbench of the HLS implementation.
The \textit{tb\_analyzer\_helper} class provides functions to analyze the data, such as comparison against defined error thresholds or the creation of graphical plots with Pythons' matplotlib library.
The \textit{tb\_data\_handler} only works as a data storage class, that encapsulates the data that is captured from the DUT.
The class called \textit{fm\_receiver\_model} serves as the 'golden model reference' that is used to compare the DUT against.
Internally, it loads the Matlab verification data.
This could be implemented differently, since Python can be used to actually implement an FM receiver including the entire DSP functionality, just like in Matlab.
Because the Matlab model already exists, this is not done here - loading the Matlab data is sufficient as it is been seen as 'the single source of truth' in this project.\\

@@

Common strategies for testing are unit tests and integration tests, which represent different levels of testing.
Unit tests specifically target a single unit, while integration tests verify the combination of multiple units as a whole.
Both of these stragegies are important.
Given all unit tests being successful, it is ensured that all these single units are implemented correctly.
However, their combination and interaction may still be subject to error - this is where integration tests unfold their potential.
Implementation errors in the glue logic between the units, as well as the interaction of the respective interfaces can be found in these integration tests.\\

@@

In this context, there are further keywords that need to be mentioned - black box testing and white box testing.
These are testing strategies, and describe how the DUT is treated.
In the case of black box, the DUTs internals, such as implementation details, are not known.
Thus, only the exposed interfaces can be tested.
In the white box as a contrary, every detail of the internal structure is known.
Thus, even single signals can be accessed, which is what is done in this Python testbench.
The advantage of the white box is, that these internals can be tested.
However, any time the implementation changes, even if it is just the naming of a signal, the testbench code needs to be adapted as well.
In black box testing, the testbench code is independent of the detailed implementation of the DUT, as long as it maintains its required behaviour at the exposed interfaces.
However, the disadvantage is that internal states cannot be observed.\\

@@

In the VHDL version of the FM radio receiver IP, this handshaking is not implemented according to the AXI stream specification.
Specifically, the ready signal handshake is implemented with a workaround, to simplify the development.
The DSP chain of the IP is not fully pipelined and thus cannot consume an input sample on every clock cycle.
This is mainly because of the FIR filter implementation - it takes multiple clock cycles to compute an output sample, before it can take a new sample.
Ideally, this situation should be handled by the ready signal, since this is exactly what it is intended for.
However, in order to simplify the implementation a strobe generator is used, which is a counter that generates a repetitive pulse with a defined period.
This pulse, in combination with some additional logic, is used as the ready signal, to guarantee enough spacing between two consecutive input samples.

@@

The VHDL code for this register interface is generated by a 'Register Engine' Python script.
The registers can be defined in a YAML-file.
Additionally, documentation in the form of a Markdown language file is generated, and - more importantly - a C/\cplusplus\ code header, which defines all the register addresses and can be used in the application software.\\

>>>>

In the implemented version, an audio file is modulated into the FM channel.
The audio file contains data that represents speech with the words \textit{'Left channel... Right channel...'}.
In terms of audio channels, the words exactly match the respective channels' output -- the first part is modulated on the left audio channel only, while the second part only targets the right audio channel.
This is particulary useful for verification, especially to verify the functionality of channel separation.
From the experience during development, the channel separation is a good indicator of whether the system is functioning correctly.
This left-right indication also simplifies the verification process, because a developer can recognize it with their visual and acoustic senses.
It is possible to audibly verify it by listening to the output.
It is also possible to just visually inspect the output data, where a clear separation is visible between the channels.\\

@@

In the final chain of DSP blocks, the sample rate is increased again, to be able to frequency-modulate the signal.
The resulting frequency modulated signal represents the signal that is received by an antenna -- the RF signal of an FM channel.
This signal is then processed by a complex baseband mixer, which results in inphase-, and quadrature components in the baseband.
An LP filter band-limits the signal and removes artifacts of the mixer.
As a final step, the sample rate is decreased again, back to the chosen sampling rate of 960~kHz.
This signal can now be processed in a digital FM receiver.

@@

Again, an audio file is used as an input, which is modulated as an FM channel signal.
The entire multiplex signal is then modulated onto an actual carrier frequency of 99~MHz, that is within the FM radio broadcast band.
This entire DSP chain is being run in the GNU Radio software, on a host PC.
Only the final signal -- FM modulated, and on a carrier of 99~MHz -- is then sent to the USRP device in the form of digital signal samples.
The USRP hardware then transmits the signal via its antenna.\\

@@

The usage of a software like GNU Radio in combination with external hardware is an interesting tool for a use-case like this project.
An FM receiver is to be implemented in multiple different versions.
All of them need to be verified somehow -- this is where the SDR transmitter created with GNU Radio and external hardware opens up an entire new possibility from the viewpoint of system level design integration.\\

@@

The testbench framework cocotb is developed as a free, open-source project.
It can be used to verify VHDL and SystemVerilog RTL code, in combination with an external simulator.
Various different simulators can be used with cocotb -- amongst others, GHDL is supported.
The interface that is used to the simulator is the industry-standard VPI.\\

@@

The Python classes \textit{tb\_analyzer\_helper}, \textit{tb\_data\_handler} and \textit{fm\_receiver\_model} are implemented to create a better structure of the testbench and to provide code that is re-useable for the testbench of the HLS implementation.
The \textit{tb\_analyzer\_helper} class provides functions to analyze the data, such as comparison against defined error thresholds or the creation of graphical plots with Pythons' matplotlib library.
The \textit{tb\_data\_handler} only works as a data storage class, that encapsulates the data that is captured from the DUT.
The class called \textit{fm\_receiver\_model} serves as the 'golden model reference' that is used to compare the DUT against.
Internally, it loads the Matlab verification data.
This could be implemented differently, since Python can be used to actually implement an FM receiver including the entire DSP functionality, just like in Matlab.
Because the Matlab model already exists, this is not done here -- loading the Matlab data is sufficient as it is been seen as 'the single source of truth' in this project.\\

@@

Common strategies for testing are unit tests and integration tests, which represent different levels of testing.
Unit tests specifically target a single unit, while integration tests verify the combination of multiple units as a whole.
Both of these stragegies are important.
Given all unit tests being successful, it is ensured that all these single units are implemented correctly.
However, their combination and interaction may still be subject to error -- this is where integration tests unfold their potential.
Implementation errors in the glue logic between the units, as well as the interaction of the respective interfaces can be found in these integration tests.\\

@@

In this context, there are further keywords that need to be mentioned -- black box testing and white box testing.
These are testing strategies, and describe how the DUT is treated.
In the case of black box, the DUTs internals, such as implementation details, are not known.
Thus, only the exposed interfaces can be tested.
In the white box as a contrary, every detail of the internal structure is known.
Thus, even single signals can be accessed, which is what is done in this Python testbench.
The advantage of the white box is, that these internals can be tested.
However, any time the implementation changes, even if it is just the naming of a signal, the testbench code needs to be adapted as well.
In black box testing, the testbench code is independent of the detailed implementation of the DUT, as long as it maintains its required behaviour at the exposed interfaces.
However, the disadvantage is that internal states cannot be observed.\\

@@

In the VHDL version of the FM radio receiver IP, this handshaking is not implemented according to the AXI stream specification.
Specifically, the ready signal handshake is implemented with a workaround, to simplify the development.
The DSP chain of the IP is not fully pipelined and thus cannot consume an input sample on every clock cycle.
This is mainly because of the FIR filter implementation -- it takes multiple clock cycles to compute an output sample, before it can take a new sample.
Ideally, this situation should be handled by the ready signal, since this is exactly what it is intended for.
However, in order to simplify the implementation a strobe generator is used, which is a counter that generates a repetitive pulse with a defined period.
This pulse, in combination with some additional logic, is used as the ready signal, to guarantee enough spacing between two consecutive input samples.

@@

The VHDL code for this register interface is generated by a 'Register Engine' Python script.
The registers can be defined in a YAML-file.
Additionally, documentation in the form of a Markdown language file is generated, and -- more importantly -- a C/\cplusplus\ code header, which defines all the register addresses and can be used in the application software.\\

### 2021-07-29 doc/thesis/chapters/introduction.tex
sample rate reduction hls
comments
results
results wip..
Fig. -> Figure
incorporate some of the feedback (wip)
fix all the 'Gedankenstriche' in sentences

Those major reasons lead to the choice of topic for this thesis - understanding all the single parts by themselves, while maintaining the understanding over an integrated embedded system that fulfills a task as a whole.

@@

\section{Objective}

@@

%develop an FM receiver and explain it, so that it can be followed in a tutorial with an affordable budget in hardware.\\

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% FEEDBACK, Meeting 07-16-2021
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%formatting:
%- absätze überall gleich (mit oder ohne leerzeile)
%references:
%- hauptsache überall gleich (Fig.2.3 oder Sec.7.1 oder Eq.3 oder Chpt.4)
%- literature refs at the beginning of a chapter/section
%  like "this chapter is based on [1][4][5]."
%  or at the end of the last sentence like "...the end of this sentence. [7]"
%
%Allgemeiner Teil / Theorie:
% zusätzliches kapitel:
%   system design ansätze
%   - from algo to VHDL (book: vom algorithmus zur hardware)
%   - matlab to VHDL
%   - systementwuf richtung test driven design
%   --> design space exploration
%
%
% ganze Arbeit aufhängen an:
%   main objective: fit all the aspects (algorithm, impl, testing) into
%                   a single project
%
%
%
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% TODO General
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% ## unify all graphs and diagrams
%       - all gray; use exact same gray tone; eliminate the yellow?
%
% ## check all abbreviation and where they are introduced the first time
%       - introduce only once, like "device under test (DUT)",
%         then use the abbreviation only
%

>>>>

Those major reasons lead to the choice of topic for this thesis -- understanding all the single parts by themselves, while maintaining the understanding over an integrated embedded system that fulfills a task as a whole.

@@

%%%%%%%%%%%%%%%%%%%%%%%
\section{Objective}

@@

%%%%%%%%%%%%%%%%%%%%%%%
\section{Outline}

TODO

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% FEEDBACK, Meeting 07-16-2021
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%formatting:
%- absätze überall gleich (mit oder ohne leerzeile)
%references:
%- hauptsache überall gleich (Fig.2.3 oder Sec.7.1 oder Eq.3 oder Chpt.4)
%- literature refs at the beginning of a chapter/section
%  like "this chapter is based on [1][4][5]."
%  or at the end of the last sentence like "...the end of this sentence. [7]"
%
%Allgemeiner Teil / Theorie:
% zusätzliches kapitel:
%   system design ansätze
%   - from algo to VHDL (book: vom algorithmus zur hardware)
%   - matlab to VHDL
%   - systementwuf richtung test driven design
%   --> design space exploration
%
%
% ganze Arbeit aufhängen an:
%   main objective: fit all the aspects (algorithm, impl, testing) into
%                   a single project
%                  --> design space exploration
%
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% FEEDBACK, Meeting 07-29-2021
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% ## Absätze
%   -- hab jetzt alle MIT Leerzeile
%
% ## Outline?
%   -- 1-2 Sätze Zusammenfassung zu jedem Chapter
%
% ## TITEL ändern
%   -- design of an fm receiver at system level using different methods
%
%   -- different approaches designing an fm receiver at system level
%
%   -- without the "fm receiver" part?
%
% ## maximale Länge der Arbeit?
%
% ## PRÜFUNG
%   -- H.G. nicht da --> 2 Möglichkeiten:  1. Oktober
%                                          2. Zweitfach PFA (AMV?)
%
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% TODO General
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% ## unify all graphs and diagrams
%       - all gray; use exact same gray tone; eliminate the yellow?
%
% ## unify all references (use format like in caption)
%       - Figure 7.10
%       - Chapter 7.3
%       - Section 7.8
%       - Program 3.5
%       - etc.
%
% ## check all abbreviation and where they are introduced the first time
%       - introduce only once, like "device under test (DUT)",
%         then use the abbreviation only
%
% ## unify highlighted text ("Betonungen")
%       - use 'asdf' or \textit{asdf}
%
%
%
%

### 2021-07-29 doc/thesis/chapters/system-design-process.tex
fix all the 'Gedankenstriche' in sentences

High-Level Synthes (HLS) is an FPGA design methodology that provides a high level of abstraction, and thus can be seen as a system level design method.
Different languages are used for HLS - two well-known languages are SystemC, which is a library for \cplusplus\ to provide modelling and simulation features, and plain \cplusplus.
A comprehensive list of available HLS tools and languages can be found in Section 3.3 of \cite{2016FfSP}.\\

>>>>

High-Level Synthes (HLS) is an FPGA design methodology that provides a high level of abstraction, and thus can be seen as a system level design method.
Different languages are used for HLS -- two well-known languages are SystemC, which is a library for \cplusplus\ to provide modelling and simulation features, and plain \cplusplus.
A comprehensive list of available HLS tools and languages can be found in Section 3.3 of \cite{2016FfSP}.\\
